Measuring contrast processing in the visual system using SSVEP (and VEP?)

### 2024-04-02 review.qmd
Rearranged

Although its definition as a stimulus is simple, the computations that underlie contrast processing in the brain have been the subject of intense research for many decades. The neural code for contrast, even in the earliest parts of visual cortex, is not simply a linear transform of the contrast at the retina - instead, contrast signals undergo a cascade of nonlinear processing stages that, broadly, attempt to normalise the output relative to the spatiotemporal environment. This normalization, achieved through a computation called 'contrast gain control' [Heeger 92, Heeger and Carandini 2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth. As an example, a grating placed at the centre of a low-contrast background tpyically appears more intense than the same grating when superimposed on a high contrast background.
```{python}
import numpy as np
import matplotlib.pyplot as plt

@@

An early use of the contrast response function was to provide an objective estimate of spatial contrast sensitivity in infants, without requiring behavioural responses.

@@

The hyperbolic ration function is monotinc but notice the slight roll-off in the amplitude of the second harmonic at high input contrasts. This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity. Power at other harmonics increases and the total power generated by the input is monotonic. This roll-off is often seen in experimental data and has been referred to as 'supersaturation' [@Tyler]

@@

>>>>

Although its definition as a stimulus is simple, the computations that underlie contrast processing in the brain have been the subject of intense research for many decades. The neural code for contrast, even in the earliest parts of visual cortex, is not simply a linear transform of the contrast at the retina - instead, contrast signals undergo a cascade of nonlinear processing stages that, broadly, attempt to normalise the output relative to the spatiotemporal environment. This normalization, achieved through a computation called 'contrast gain control' [Nachmas and Sansbury? Foley? Heeger 92, Heeger and Carandini 2011] maximises the sensitivity of the visual system by making optimal use of neuronal bandwidth. As an example, a grating placed at the centre of a low-contrast background tpyically appears more intense than the same grating when superimposed on a high contrast background.
```{python}
import numpy as np
import matplotlib.pyplot as plt

@@

To understand the utility of the contrast SSVEP, it is helpful to identify the cascade of processing stages in the early visual system that give rise to it. In the following section we illustrate how a typical SSVEP signal measured over early visual cortex might contain information about a large number of early visual computations.

@@

The hyperbolic ration function is monotonic but notice the slight roll-off in the amplitude of the second harmonic at high input contrasts. This results from the distortion of the input sine waves at high contrast due to a combination of the full-wave rectification and saturating non-linearity. Power at other harmonics increases and the total power generated by the input is monotonic. This roll-off is often seen in experimental data and has been referred to as 'supersaturation' [@Tyler]

@@

An early use of the SSVEP was to provide an objective estimate of spatial contrast sensitivity in infants, without requiring behavioural responses.
